---
title: "p8105_hw3_sw3737"
author: "sw3737"
date: "2023-10-14"
output: github_document
---

## Problem 2

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
```

# Data import

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

# Data cleaning

```{r}
brfss_clean <- brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health", 
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

# State analysis for 2002 and 2010

```{r}
brfss_clean %>%
  group_by(year, locationabbr) %>%
  summarise(counties = n_distinct(locationdesc)) %>%
  filter((year == 2002 | year == 2010) & counties >= 7)
```

Comments: In 2002, 6 states were observed at 7 or more locations (CT, FL, MA, NC, NJ, PA). In 2010, 14 states were observed at 7 or more locations (CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA).

# Construct a dataset

```{r}
excellent_data <- brfss_clean %>% 
  filter(response == "Excellent") %>%
  group_by(year, locationabbr) %>%
  summarise(avg_data_value = mean(data_value, na.rm = TRUE), .groups = "drop")

ggplot(excellent_data, aes(x = year, y = avg_data_value, group = locationabbr, color = locationabbr)) +
  geom_line() +
  labs(title = "Average Value of 'Excellent' Responses Over Time by State",
       x = "Year", y = "Average Value") +
  theme_minimal() 
```

Comments: The "spaghetti" plot illustrates the average value of "Excellent" responses over time for each state, spanning the years 2002 to 2010. The graph is populated with a multitude of colorful lines, each representing a different state, showcasing the variation in responses over the years. Many states exhibit fluctuating trends, indicating changes in the perception of what constitutes an "Excellent" response. 

# Two-Panel Plot for 2006 and 2010

```{r}
ny_data <- brfss_clean %>% 
  filter(locationabbr == "NY", year %in% c(2006, 2010))

ggplot(ny_data, aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_wrap(~ year) +
  labs(title = "Distribution of Responses in NY for 2006 and 2010",
       x = "Response", y = "Data Value") +
  theme_minimal()
```

Comments: These two-panel plots show the distribution of `data_value` for responses from `poor` to `excellent` among locations in NY State. In 2006, the median value for `Good` and `Very good` responses is relatively high, while by 2010, the median for these categories has slightly decreased. Conversely, the `Excellent` category saw a rise in its median value from 2006 to 2010. This suggests a shift in the distribution of responses over this four-year period in New York.

## Problem 3

```{r set, include=FALSE}
library(readr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(janitor)
```

# Demographics data cleaning

```{r}
nhanes_covar <- read_csv("Data files/nhanes_covar.csv", skip = 4) %>%
  janitor::clean_names() %>% 
  mutate(
    sex = recode(sex, "1" = "male", "2" = "female"),
    education = recode(education, 
                       "1" = "Less than high school",
                       "2" = "High school equivalent",
                       "3" = "More than high school")
  ) %>%
  filter(age >= 21 & 
         !is.na(sex) & 
         !is.na(age) & 
         !is.na(bmi) & 
         !is.na(education))
```

# Accelerometer data cleaning and merge the two datasets

```{r}
nhanes_accel <- read_csv("Data files/nhanes_accel.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    min1:min1440,
    names_to = "min",
    values_to = "min_value"
  ) %>%
  separate(min, into = c("prefix","min"), sep = 3) %>%
  select(-prefix)
  
merged_data <- inner_join(nhanes_accel, nhanes_covar, by = "seqn") %>%
    mutate(education = factor(education, levels = c("Less than high school", "High school equivalent", "More than high school")))
```

# Table of number of men and women in each education category
```{r}
table_data <- merged_data %>% 
  group_by(education, sex) %>% 
  summarise(count = n())

table_data
```

Comments: It displays a tabulated count of individuals based on their education level and gender. It highlights three education categories: `Less than high school`, `High school equivalent`, and `More than high school`. For `Less than high school` and `More than high school level`, females tend to have slightly higher counts than males.

# Visualization of age distributions for men and women in each education category
```{r}
ggplot(merged_data, aes(x = age, fill = sex)) +
  geom_histogram(alpha = 0.7, position = "identity") +
  facet_wrap(~education) +
  labs(title = "Age Distribution by Sex and Education",
       x = "Age",
       y = "Count") +
  theme_minimal() +
  scale_fill_manual(values = c("male" = "blue", "female" = "pink"))
```

Comments: It illustrates the age distribution of individuals categorized by sex and education. It's observed that for the `Less than high school` education category, the `highest counts are concentrated around the age range of 60-80 years`. While for the `More than high school` category, there's a `significant rise in counts for females between the ages of 20-40`. The chart provides a visual representation of how age and educational attainment intersect with gender.

# Plot of total activities vs age
```{r}
total_activity_data <- merged_data %>%
  group_by(seqn, age, sex, education) %>%
  summarise(total_activity = sum(min_value, na.rm = TRUE))

ggplot(total_activity_data, aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~education) +
  labs(title = "Total Activity vs Age by Sex and Education",
       x = "Age",
       y = "Total Activity") +
  theme_minimal() +
  scale_color_manual(values = c("male" = "blue", "female" = "pink"))
```

Comments: There is a `declining trend in total activity with increasing age across all education levels`. In all three education categories, `females generally have a higher total activity level` than males across various age groups. The decline in activity with age seems steeper for those with `Less than high school` education as compared to the other two categories.

# 24-Hour activity time courses plot
```{r}
activity_time_course <- merged_data %>%
  group_by(min, education, sex) %>%
  summarise(avg_activity = mean(min_value, na.rm = TRUE))

ggplot(activity_time_course, aes(x = as.numeric(min), y = avg_activity, color = sex)) +
  geom_line() +
  facet_wrap(~education) +
  labs(title = "24-Hour Activity Time Course by Sex and Education",
       x = "Time (Minutes from Midnight)",
       y = "Average Activity") +
  theme_minimal() +
  scale_color_manual(values = c("male" = "blue", "female" = "pink"))
```

Comments: The graph seems to follow a `diurnal rhythm`, where there's an increase in activity during the day and a decrease during the nighttime for all education levels. The peaks of activity occur approximately at mid-day. In all three education categories, `females tend to have higher activity levels during the day` as compared to males.There is very `minimal activity during the late-night hours` for all groups, indicating that most participants are likely at rest.